---
title: "simpleRegress"
author: "Peter Dang"
date: "Spring 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

According to the observation for WEIGHT explained by NECK, we can see that the RSE is 43.91 and the adjusted R-squared is 0.87.  

## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
mHEADLEN <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(mHEADLEN)
summary.lm(mHEADLEN)
```

According to the observations, the residual standard error is 67.81, and our adjusted R-squared is 0.69. Ordering the models from best to worst, the best one is the weight explained by neck followed by the weight explained by HEADLEN. 

NECK RSE is 43.91  
HEADLEN RSE is 67.81  

```{r}
mHEADWTH <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(mHEADWTH)
summary.lm(mHEADWTH)
```

We can see that for this plot, the residual standard error is 76.41 and the adjusted R squared is 0.6064. 

Best to worst ordered by RSE    
NECK RSE is 43.91     
HEADLEN RSE is 67.81       
HEADWTH RSE is 76.41   

```{r}
mLENGTH <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(mLENGTH)
summary.lm(mLENGTH)
```

For this plot, we can see that the RSE is 61.83 and the adjusted R-squared is 0.7423.  
Best to worst ordered by RSE      
NECK RSE is 43.91      
LENGTH RSE is 61.83     
HEADLEN RSE is 67.81         
HEADWTH RSE is 76.41  

